---
title: "Assignment 5 - Meta-analysis of pitch in schizophrenia"
author: "Riccardo Fusaroli"
date: "3/7/2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Building on the shoulders of giants: meta-analysis

## Questions to be answered

1. What is the current evidence for distinctive patterns of pitch mean and pitch sd in schizophrenia? Report how many papers report quantitative estimates, your method to analyze them, the estimated effect size of the difference (mean effect size and standard error for pitch mean, same for pitch sd) and forest plots representing it. 
```{r setup, include=FALSE}
setwd("~/OneDrive/3 Semester/Experimental Methods 3/Assignment 5")
data = read.csv("Data.csv", sep = ";")

library(pacman) 
library(lmerTest)
p_load(metafor)

#Calculating Cohen's d (yi) and the variance of Cohen's d (vi). Adding the values to the data
PitchMean=escalc('SMD', n1i=SampleSizeSchizo, n2i=SampleSizeContros, m1i=PitchMeanSchizo, m2i=PitchMeanControls, sd1i=PitchMeanSchizoSD, sd2i=PitchMeanControlsSD, data = data)

#Doing an lmer on the studies with Cohen's d, only four studies
m1 <- lmer(yi ~ 1 + (1 | Article), weights = 1/vi, data=PitchMean, control=lmerControl(check.nobs.vs.nlev="ignore", check.nobs.vs.nRE="ignore"))
summary(m1)

#Meta-analysis optimization: 
res <- rma(yi, vi, data = PitchMean, slab=Article) 
forest(res)

#Calcuclating the number of participants in the studies used
PitchMean2 = subset(PitchMean[PitchMean$vi > 0,])
sum(PitchMean2$SampleSizeSchizo, na.rm = TRUE)
sum(PitchMean2$SampleSizeContros, na.rm = TRUE)

#Repeating the procedure for the SD of the SDs. 
PitchSD=escalc('SMD', n1i=SampleSizeSchizo, n2i=SampleSizeContros, m1i=PitchSDSchizo, m2i=PitchSDControls, sd1i=PitchSDSchizoSD, sd2i=PitchSDControlsSD, data = data)

m2 <- lmer(yi ~ 1 + (1 | Article), weights = 1/vi, data=PitchSD, control=lmerControl(check.nobs.vs.nlev="ignore", check.nobs.vs.nRE="ignore"))
summary(m2)

#(beta = -0.125, SE = 0.579, t() = -0.216)
#Meta-analysis optimization: 
res2 <- rma(yi, vi, data = PitchSD, slab=Article) 
forest(res2)

#Calcuclating the number of participants in the studies used
PitchSD2 = subset(PitchSD[PitchSD$vi > 0,])
sum(PitchSD2$SampleSizeSchizo, na.rm = TRUE)
sum(PitchSD2$SampleSizeContros, na.rm = TRUE)

```


2. Do the results match your own analysis from Assignment 3? If you add your results to the meta-analysis, do the estimated effect sizes change? Report the new estimates and the new forest plots.

```{r setup, include=FALSE}
own_data = read.csv("Schizophrenia_data.csv")
own_data = own_data[-1:-2]

PitchMeanSchizo = mean(own_data$mean[own_data$Diagnosis == "Schizophrenia"])
PitchMeanControls = mean(own_data$mean[own_data$Diagnosis == "Control"])

PitchSDSchizo = mean(own_data$SD[own_data$Diagnosis == "Schizophrenia"])
PitchSDControls = mean(own_data$SD[own_data$Diagnosis == "Control"])

PitchMeanSchizoSD = sd(own_data$mean[own_data$Diagnosis == "Schizophrenia"])
PitchMeanControlsSD = sd(own_data$mean[own_data$Diagnosis == "Control"])

PitchSDSchizoSD = sd(own_data$SD[own_data$Diagnosis == "Schizophrenia"])
PitchSDControlsSD = sd(own_data$SD[own_data$Diagnosis == "Control"])

SampleSizeSchizo = as.numeric(length(unique(own_data$Subject[own_data$Diagnosis == "Schizophrenia"])))
SampleSizeContros = as.numeric(length(unique(own_data$Subject[own_data$Diagnosis == "Control"])))
Article = "Own_Analysis"
Year = 2017
own_analysis = data.frame(Article, Year, SampleSizeSchizo, SampleSizeContros, PitchMeanControls, PitchMeanControlsSD, PitchMeanSchizo,  PitchMeanSchizoSD, PitchSDControls, PitchSDControlsSD, PitchSDSchizo, PitchSDSchizoSD)

all_data = rbind(own_analysis,data)


#Running the analysis again, now with our own study as well.
#Calculating Cohen's d (yi) and the variance of Cohen's d (vi). Adding the values to the data
EffectsizeMean=escalc('SMD', n1i=SampleSizeSchizo, n2i=SampleSizeContros, m1i=PitchMeanSchizo, m2i=PitchMeanControls, sd1i=PitchMeanSchizoSD, sd2i=PitchMeanControlsSD, data = all_data)

#Doing an lmer on the studies with Cohen's d, only four studies
m3 <- lmer(yi ~ 1 + (1 | Article), weights = 1/vi, data=EffectsizeMean, control=lmerControl(check.nobs.vs.nlev="ignore", check.nobs.vs.nRE="ignore"))
summary(m3)

#Meta-analysis optimization: 
res3 <- rma(yi, vi, data = EffectsizeMean, slab=Article) 
forest(res3)

#Calcuclating the number of participants in the studies used
EffectsizeMean2 = subset(EffectsizeMean[EffectsizeMean$vi > 0,])
sum(EffectsizeMean2$SampleSizeSchizo, na.rm = TRUE)
sum(EffectsizeMean2$SampleSizeContros, na.rm = TRUE)

#Repeating the procedure for the SD of the SDs. 
EffectsizeSD=escalc('SMD', n1i=SampleSizeSchizo, n2i=SampleSizeContros, m1i=PitchSDSchizo, m2i=PitchSDControls, sd1i=PitchSDSchizoSD, sd2i=PitchSDControlsSD, data = all_data)

m4 <- lmer(yi ~ 1 + (1 | Article), weights = 1/vi, data=EffectsizeSD, control=lmerControl(check.nobs.vs.nlev="ignore", check.nobs.vs.nRE="ignore"))
summary(m4)

#Meta-analysis optimization: 
res4 <- rma(yi, vi, data = EffectsizeSD, slab=Article) 
forest(res4)

#Calcuclating the number of participants in the studies used
EffectsizeSD2 = subset(EffectsizeSD[EffectsizeSD$vi > 0,])
sum(EffectsizeSD2$SampleSizeSchizo, na.rm = TRUE)
sum(EffectsizeSD2$SampleSizeContros, na.rm = TRUE)

```


3. Assess the quality of the literature: report and comment on heterogeneity of the studies (tau, I2), on publication bias (funnel plot), and on influential studies.

```{r setup, include=FALSE}
summary(res3)
# tau^2 (estimated amount of total heterogeneity) = 0.021, SE = 0.055. I^2 (total heterogeneity/total variability) = 26.15 %. The higher I^2 value the higher the percentage is that they come from different population. 

summary(res4)
# tau^2 (estimated amount of total heterogeneity) = 3.499, SE = 2.07. I^2 (total heterogeneity/total variability) = 98.18 %. The higher I^2 value the higher the percentage is that they come from different population. The high value is due to Cohen et al. 2014. 

funnel(res3, main = "Random-Effects Model",xlab = "Standardized Mean Difference")
regtest(res3)
ranktest(res3)

funnel(res4, main = "Random-Effects Model",xlab = "Standardized Mean Difference")
regtest(res4)
ranktest(res4)


#Testing influential studies:
inf3 <- influence(res3) 
print(inf3) 
plot(inf3)


inf4 <- influence(res4) 
print(inf4) 
plot(inf4)


```

#When you have model1 you can do summary(model1), tau^2 - tells you how spread the area covered by the plots are, the variance. I^2 - the percentage of tau^2 that could not be explained by uncertainty (if the data is outside the confidence interval (if studies do not overlap in the forest plot)). More on Tuesday

## Tips on the process to follow:

- Download the data on all published articles analyzing pitch in schizophrenia: https://www.dropbox.com/s/pmgw8wzc308so1p/Data.csv?dl=0
- Following the procedure in the slides calculate effect size and standard error of the effect size per each study. N.B. we focus on pitch mean and pitch standard deviation.
 . first try using lmer (to connect to what you know of mixed effects models)
 . then use rma() (to get some juicy additional statistics)

- Build a forest plot of the results (forest(model))
 
- Go back to Assignment 3, add your own study to the data table, and re-run meta-analysis. Do the results change?

- Now look at the output of rma() and check tau and I2
